---
title: "607-Group-Project-3"
author: "Steve Phillips, Dirk Hartog, Lwin Shwe"
date: "2023-10-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Probem Statement


This is a project for all the members of an assigned group to work on together, since being able to work effectively on a virtual team is a key “soft skill” for data scientists. 
Use data to answer the question, “Which are the most valued data science skills?”
Please note especially the requirement about making a presentation during the class meetup on the following Wednesday.

Effective communication skill is essential for data scientists, and the ability to communicate the findings to team and explain what value these insights can bring to technical or non-technical stakeholders also are very important. 
Curiosity is a trait that makes data science more interesting and beneficial. Being curious about business problems will let you to the solution as well.
Team works in a team setting and being a team player is critical. We should always listen to their team members for their input and take into consideration to work as a team.


```{r load-package}
library(RCurl)
library(kableExtra)
library(DBI)
library(RMySQL)
library(RODBC)
conn <- dbConnect(MySQL(), user= "root", password = "M2max2023", dbname='MyDB1', host='localhost', port=3306)
```

### Load realational database from MySQL Workbench to R

```{r DBMS}
library(dplyr)
library(ggplot2)
cryto_dbms <- dbGetQuery(conn , "select theDate,opening, opening, highest, lowest, closing, volume, marketcap, DailyPrice.CoinSymbol, coalesce(CoinDesc,'') CoinDesc
from DailyPrice
left join CoinList
on DailyPrice.CoinSymbol = CoinList.CoinSymbol")
glimpse(cryto_dbms)
# Export the data frame to a CSV file local directory
write.csv(cryto_dbms, "DATA 607 Group Project3.csv", row.names = FALSE)
#Read the .csv file 
cryptocurrencies <- read.csv("DATA 607 Group Project3.csv")
head(cryptocurrencies)

```
```{r monthly-volume}
total_volume_in_month <- dbGetQuery(conn, "select theMonth, CoinSymbol, sum(volume) TotalVolumeInMonth
from
(
select monthname(theDate) theMonth, CoinSymbol, volume 
from DailyPrice
where year(theDate) = 2021 and CoinSymbol in ('ADA', 'ETH') 
) A
group by theMonth, CoinSymbol;")
total_volume_in_month

```
### What would be the total monthly volume of each coin in the year 2023?
 
```{r}
cryptocurrencies %>%
count(CoinSymbol) # count number of coins types 
summary(cryto_dbms)

filter(cryptocurrencies,CoinSymbol == "ADA", theDate == "2023-09-13")


# Create the line graph
ggplot(cryptocurrencies, aes(x = CoinSymbol, y = volume, group = CoinSymbol)) +
  geom_line() +
  labs(x = "CoinSymbol", y = "volume") +
  theme_minimal()


```
```{r cleaning}

# Checking for missing values
missing_values <- sapply(cryptocurrencies, function(x) sum(is.na(x)))

# Display the columns with missing values
columns_with_missing <- names(missing_values[missing_values > 0])
print(columns_with_missing)

# Display the number of missing values for each column
print(missing_values)

# Create box plots for numeric columns that you want to check
numeric_columns <- cryptocurrencies %>%
  select(opening, highest, lowest, closing, volume, marketcap)

# Create a box plot for each numeric column
for (col in names(numeric_columns)) {
  p <- ggplot(cryptocurrencies, aes(y = !!sym(col))) +
    geom_boxplot() +
    labs(title = paste("Box Plot of", col))
  print(p)
}

```

```{r transformation}
#Standardizing Dates
# Eg. Converting a date column to a standard format
cryptocurrencies$theDate <- as.Date(cryptocurrencies$theDate, format = "%Y-%m-%d")
head(cryptocurrencies$theDate)

#Scaling Numerical Values
# Eg. Scaling the "volume" column
cryptocurrencies$volume <- scale(cryptocurrencies$volume)
head(cryptocurrencies$volume)

#Aggregating Data using (group_by)
# Example: Aggregating data by CryptoDesc and calculating the mean marketCap
aggregated_data <- cryptocurrencies %>%
  group_by(CoinSymbol) %>%
  summarize(mean_marketcap = mean(marketcap))
aggregated_data

```
